---
title: "THP1 cell tracking + characterization"
author: "Marion Hardy"
date: Sys.Date()
output: 
  html_document: 
    toc: true
    theme: spacelab
    highlight: monochrome
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, cache = TRUE, echo = FALSE, warning = F, cache.lazy = F)
knitr::opts_chunk$set(fig.width=10, fig.height=12) 

library(tidyverse)
library(XML)
library(openxlsx)
library(cowplot)
library(ggpubr)
library(rstatix)
library(kableExtra)

```

```{r Data loading, include=FALSE}

xys = c("xy01","xy02","xy05","xy06","xy07","xy08","xy12","xy13","xy14","xy15",
        "xy18","xy19","xy20","xy21","xy24","xy25")


for(i in 1:length(xys)){
  tracks = read.csv(paste0("./data_output/",xys[i],"/",
                           paste0(xys[i],"_spots.csv")), stringsAsFactors = F)
  tracks = tracks[-c(1:3),]
  tracks[,-1] = as.numeric(unlist(tracks[,-1])) # because of the first 3 lines being chr, gotta change table into num
  tracks$TRACK_ID = as.factor(tracks$TRACK_ID)
  tracks$xy = xys[i]
  tracks = tracks %>% filter(!is.na(TRACK_ID))
  tracks = arrange(tracks, POSITION_T) # to get geom_path right
  assign(xys[i],tracks)
}


```

# Introduction

```{r}

head(tracks)

head(table(tracks$TRACK_ID, tracks$FRAME))

```

TRACK_ID is the reference to the track, the number of time it appears is the number of frames it was tracked through. 
Every line is the track (center of the spot), its frame and the characteristics of the spot being tracked (area, circularity, intensity, ...)

# Spider plotting the tracks if possible

```{r, fig.height=3, fig.width=3}

p1 = 
xy02 %>% 
  ggplot(aes(x = POSITION_X,
             y = POSITION_Y,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "xy02: 1000 cells, glass")

# little bit of a mindfuck but these are flipped horizontally compared to the
# imageJ trackmate video output

p2 = 
xy05 %>% 
  ggplot(aes(x = POSITION_X,
             y = POSITION_Y,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "xy05: 1000 cells, collagen")

p3 = 
xy12 %>% 
  ggplot(aes(x = POSITION_X,
             y = POSITION_Y,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "xy12: 2000 cells, glass")

p4 = 
xy07 %>% 
  ggplot(aes(x = POSITION_X,
             y = POSITION_Y,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "xy07: 2000 cells, collagen")

p5 = 
xy14 %>% 
  ggplot(aes(x = POSITION_X,
             y = POSITION_Y,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "xy14: 5000 cells, glass")

p6 = 
xy18 %>% 
  ggplot(aes(x = POSITION_X,
             y = POSITION_Y,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "xy18: 5000 cells, collagen")

p7 = 
xy24 %>% 
  ggplot(aes(x = POSITION_X,
             y = POSITION_Y,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "xy24: 10 000 cells, glass")

p8 = 
xy20 %>% 
  ggplot(aes(x = POSITION_X,
             y = POSITION_Y,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "xy20: 10 000 cells, collagen")

```

```{r, fig.height=8, fig.width=16}

plot_grid(p1, p2, p3, p4, p5, p6, p7, p8, ncol = 4 , align = "hv")

```

## Centering the tracks to 0,0


```{r}

vars = list(xy01,xy02,xy05,xy06,xy07,xy08,xy12,xy13,xy14,xy15,
        xy18,xy19,xy20,xy21,xy24,xy25)

for(i in 1:length(vars)){
  vars_0 =
  vars[[i]] %>% 
  group_by(TRACK_ID) %>% 
  filter(FRAME == min(FRAME)) %>% 
  select(TRACK_ID, POSITION_X, POSITION_Y)
  
  colnames(vars_0) = c("TRACK_ID", "X_initial", "Y_initial")
  upd_xy = left_join(vars[[i]], vars_0)

  upd_xy = 
  upd_xy %>% 
    group_by(TRACK_ID) %>% 
    mutate(X_adjusted = POSITION_X - X_initial,
         Y_adjusted = POSITION_Y - Y_initial)
  assign(xys[i],upd_xy)}

```



```{r}

p1 = 
xy02 %>% 
  ggplot(aes(x = X_adjusted,
             y = Y_adjusted,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  scale_x_continuous(limits = c(-500, 500))+
  scale_y_continuous(limits = c(-500, 500))+
  theme(legend.position = "none")+
  labs(title = "xy02: 1000 cells, glass")

# little bit of a mindfuck but these are flipped horizontally compared to the
# imageJ trackmate video output

p2 = 
xy05 %>% 
  ggplot(aes(x = X_adjusted,
             y = Y_adjusted,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  scale_x_continuous(limits = c(-500, 500))+
  scale_y_continuous(limits = c(-500, 500))+
  theme(legend.position = "none")+
  labs(title = "xy05: 1000 cells, collagen")

p3 = 
xy12 %>% 
  ggplot(aes(x = X_adjusted,
             y = Y_adjusted,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  scale_x_continuous(limits = c(-500, 500))+
  scale_y_continuous(limits = c(-500, 500))+
  theme(legend.position = "none")+
  labs(title = "xy12: 2000 cells, glass")

p4 = 
xy07 %>% 
  ggplot(aes(x = X_adjusted,
             y = Y_adjusted,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  scale_x_continuous(limits = c(-500, 500))+
  scale_y_continuous(limits = c(-500, 500))+
  theme(legend.position = "none")+
  labs(title = "xy07: 2000 cells, collagen")

p5 = 
xy14 %>% 
  ggplot(aes(x = X_adjusted,
             y = Y_adjusted,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  scale_x_continuous(limits = c(-500, 500))+
  scale_y_continuous(limits = c(-500, 500))+
  theme(legend.position = "none")+
  labs(title = "xy14: 5000 cells, glass")

p6 = 
xy18 %>% 
  ggplot(aes(x = X_adjusted,
             y = Y_adjusted,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  scale_x_continuous(limits = c(-500, 500))+
  scale_y_continuous(limits = c(-500, 500))+
  theme(legend.position = "none")+
  labs(title = "xy18: 5000 cells, collagen")

p7 = 
xy24 %>% 
  ggplot(aes(x = X_adjusted,
             y = Y_adjusted,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  scale_x_continuous(limits = c(-500, 500))+
  scale_y_continuous(limits = c(-500, 500))+
  theme(legend.position = "none")+
  labs(title = "xy24: 10 000 cells, glass")

p8 = 
xy20 %>% 
  ggplot(aes(x = X_adjusted,
             y = Y_adjusted,
             color = TRACK_ID))+
  geom_path()+ 
  theme_bw()+
  scale_x_continuous(limits = c(-500, 500))+
  scale_y_continuous(limits = c(-500, 500))+
  theme(legend.position = "none")+
  labs(title = "xy20: 10 000 cells, collagen")


```



```{r, fig.height=8, fig.width=16}

plot_grid(p1, p2, p3, p4, p5, p6, p7, p8, ncol = 4 , align = "hv")

```

# Let's get some numbers on that

## Checking average displacement

```{r, fig.width=6, fig.height=3}

# This time i need to import the tracks table with their characteristics


for(i in 1:length(xys)){
  tracks = read.csv(paste0("./data_output/",xys[i],"/",
                           paste0(xys[i],"_tracks.csv")), stringsAsFactors = F)
  tracks = tracks[-c(1:3),]
  tracks[,-1] = as.numeric(unlist(tracks[,-1])) # because of the first 3 lines being chr, gotta change table into num
  tracks$TRACK_ID = as.factor(tracks$TRACK_ID)
  tracks$xy = xys[i]
  tracks = tracks %>% filter(!is.na(TRACK_ID))
  assign(xys[i],tracks)
}

vars = list(xy01,xy02,xy05,xy06,xy07,xy08,xy12,xy13,xy14,xy15,
        xy18,xy19,xy20,xy21,xy24,xy25)

# merge

all = vars %>% 
  purrr::reduce(full_join)

```


```{r}
# Annotate the data further

all =
  all %>% 
  mutate(Surface = ifelse(
    xy %in% c("xy01","xy02","xy12","xy13","xy14","xy15",
        "xy24","xy25"), "Glass","Collagen")) %>% 
  mutate(Cell_seeded = ifelse(
    xy %in% c("xy01","xy02","xy05","xy06"), "1000 cells",
    ifelse(xy %in% c("xy12","xy13","xy07","xy08"), "2000 cells",
    ifelse(xy %in% c("xy14","xy15","xy18","xy19"), "5000 cells",
    ifelse(xy %in% c("xy24","xy25","xy20","xy21"), "10 000 cells",NA)))))


table(all$Cell_seeded, all$Surface)


```

```{r}

p1 = 
all %>% 
  ggplot(aes(x = TRACK_DISPLACEMENT, col = Surface))+
  geom_freqpoly()+
  theme_bw()

p2 = 
all %>% 
  ggplot(aes(x = log(TRACK_DISPLACEMENT), col = Surface))+
  geom_freqpoly()+
  theme_bw()
p3 = 
all %>% 
  ggplot(aes(x = TOTAL_DISTANCE_TRAVELED, col = Surface))+
  geom_freqpoly()+
  theme_bw()

p4 = 
all %>% 
  ggplot(aes(x = log(TOTAL_DISTANCE_TRAVELED), col = Surface))+
  geom_freqpoly()+
  theme_bw()


```   

```{r, fig.height=8, fig.width=12}

plot_grid(p1, p2, p3, p4, ncol = 2 , align = "hv")

```


## Get some stats

### I'm assuming my data does not follow a normal distribution: Wilcoxon test

Maybe the appropriate test would be a Mann-Whitney if we consider the samples independent

```{r}

## get some values about tracks

all %>%
  group_by(Surface) %>% 
  select(TRACK_DURATION, TRACK_MEAN_SPEED,TRACK_MEDIAN_SPEED, 
         TOTAL_DISTANCE_TRAVELED, MEAN_DIRECTIONAL_CHANGE_RATE,
         LINEARITY_OF_FORWARD_PROGRESSION, TRACK_DISPLACEMENT) %>% 
  summarise(across(everything(), list(mean = mean, median = median))) %>% 
  t() %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


## stat test 

stat.disp <- all %>% wilcox_test(TRACK_DISPLACEMENT ~ Surface, 
                                 ref.group = "Glass",
                                 exact = FALSE, alternative = "less") %>%
  add_significance()


stat.tot <- all %>% wilcox_test(TOTAL_DISTANCE_TRAVELED ~ Surface, 
                                ref.group = "Glass",
                                 exact = FALSE, alternative = "less") %>%
  add_significance()


stat.lin <- all %>% wilcox_test(LINEARITY_OF_FORWARD_PROGRESSION ~ Surface, 
                                ref.group = "Glass",
                                 exact = FALSE, alternative = "less") %>%
  add_significance()


stat.dir <- all %>% wilcox_test(MEAN_DIRECTIONAL_CHANGE_RATE ~ Surface, 
                                ref.group = "Glass",
                                 exact = FALSE, alternative = "less") %>%
  add_significance()


stat.sp <- all %>% wilcox_test(TRACK_MEAN_SPEED ~ Surface, 
                               ref.group = "Glass",
                               exact = FALSE, alternative = "less") %>%
  add_significance()


# merge stats

stats = list(stat.dir, stat.disp,stat.lin,stat.sp,stat.tot)

all_stats = stats %>% 
  purrr::reduce(full_join)

all_stats %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```


```{r}

## track displacement

p1 = 
all %>% 
  ggplot(aes(x = xy, y = TRACK_DISPLACEMENT, fill = Surface))+
  geom_violin()+
  geom_jitter(color="black", size=0.7, alpha=0.25) +
  theme_bw()+
  labs(title = "No filtering")

p2 = 
all %>% 
  filter(TRACK_DISPLACEMENT > 25) %>% 
  ggplot(aes(x = xy, y = TRACK_DISPLACEMENT, fill = Surface))+
  geom_boxplot()+
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_bw()+
  labs(title = " >25 pixels ")

## track total distance traveled

p3 = 
all %>% 
  ggplot(aes(x = xy, y = TOTAL_DISTANCE_TRAVELED, fill = Surface))+
  geom_violin()+
  geom_jitter(color="black", size=0.7, alpha=0.25) +
  theme_bw()+
  labs(title = "No filtering")

p4 = 
all %>% 
  filter(TRACK_DISPLACEMENT > 25) %>% 
  ggplot(aes(x = xy, y = TOTAL_DISTANCE_TRAVELED, fill = Surface))+
  geom_boxplot()+
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_bw()+
  labs(title = " >25 pixels ")

## linearity of travel

p5 = 
all %>% 
  ggplot(aes(x = xy, y = LINEARITY_OF_FORWARD_PROGRESSION, fill = Surface))+
  geom_violin()+
  geom_jitter(color="black", size=0.7, alpha=0.25) +
  theme_bw()+
  labs(title = "Linearity of travel")

## track mean speed

p6 = 
all %>% 
  ggplot(aes(x = xy, y = MEAN_DIRECTIONAL_CHANGE_RATE, fill = Surface))+
  geom_violin()+
  geom_jitter(color="black", size=0.7, alpha=0.25) +
  theme_bw()+
  labs(title = "Mean directional change")

```

### Violin plots of motion parameters

```{r, fig.width=15, fig.height=10}

plot_grid(p1, p2, p3, p4, p5, p6, ncol = 2 , align = "hv")

```

```{r, fig.width=6, fig.height=4}
p7 = 
all %>% 
  ggplot(aes(x = xy, y = TRACK_MEAN_SPEED, fill = Surface))+
  geom_violin()+
  geom_jitter(color="black", size=0.7, alpha=0.25) +
  theme_bw()+
  labs(title = "Mean speed")

p7
```


-   TRACK_DISPLACEMENT: length of the vector between first frame point coordinates and last frame point coordinates

-   TOTAL_DISTANCE_TRAVELED: total length of cell track

-   LINEARITY OF FORWARD PROGRESSION : Linearity of travel = mean straight line speed /track mean speed.

-   MEAN_DIRECTIONAL_CHANGE_RATE :measures the angle between two succeeding links, averaged over all the links of a track



# Conclusions



# R session info

```{r}
sessionInfo()
```
